### 2017-09-29 manuscripts/AmNat_response_to_rev.Rmd
WIP

--- 
title: "Response to reviewers"
author:
  - "Milad Memarzadeh"
  - "Carl Boettiger"
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
bibliography: citations.bib
---

ASSOCIATE EDITOR'S RECOMMENDATION

Dear Professor Bronstein; Prof Dwyer,

> My recommendation is that manuscript 57722 be declined without prejudice.   The authors use a state of the art computer algorithm known as ``POMDP'' to determine, on the fly, how fisheries should be managed, in terms of the total amount of stock that can be harvested, and the profit accruing from that harvest.   This is an important problem in ecology, and the authors approach appears to offer significant insight.  The reviewers nevertheless point out some important ways in which the text could be improved, and reviewer 1 in particular argues that the authors's approach may be significantly flawed.  These problems would therefore need to be addressed for a revision to be successful.

We appreciate the careful and constructive comments of the editor and reviewers, which have allowed us to re-write the manuscript to both better communicate both the method and broader implications of the results shown here.  We also now address directly the significant differences between the extensive literature on the optimal management of fisheries to

We appreciate the time and effort of reviewer 1 in exploring these results in further simulation, but were unable to understand precisely what the reviewer had implemented. From the description provided, it appears the reviewer's results rather underscore our point about the importance of observation uncertainty, an uncertainty which does not appear to be fully present in the reviewer's simulation, as we discuss below.  In short, if the stock size is unknown (due to measurement uncertainty), how does one enforce the safe-guard that total allowable catch quota or harvest not exceed 90% the stock?  As our work shows, it precisely the presence of such assumptions

We have expanded our results to include a comparisons to policies that much more closely resemble the implementation in actual fisheries

and believe our revisions should now address each of the concerns raised.  In particular,

> One of Reviewer 2's main comments is that the article focusses too narrowly on fisheries and stock assessment, and the Reviewer argues that the authors could instead frame their work more broadly in terms of applied ecology in general, perhaps by citing problems of measurement uncertainty in other applications besides fisheries.  These points are very well taken, especially given that The American Naturalist covers biological problems in a very broad sense.   For the work to be of general interest to the readership of The American Naturalist, the authors would therefore have to explain more clearly the relevance of their work to applied ecology in general.

we now emphasize that making decisions in face of imperfect information is by no means an issue isolated solely to fisheries, and that our focus on fisheries example is merely for the same reason introductory texts on ecological management have so often turned to fisheries: no other example of ecological managment has a rich or foundational a history in our understanding of ecosystem management and decision-theory.  We now discuss the broader concept of uncertain decision making in the introduction and highlight examples where these issues most readily appear in so many other ecological applications.

> Reviewer 1 in contrast makes a  specific and detailed comment about a very particular issue.   In discussing this comment, an important point is that I am sufficiently far from being a fisheries biologist that I am not entirely understand sure that I understand the objection, emphasizing the importance of Reviewer 2's comment that The American Naturalist is a general interest journal, not a fisheries journal.  This caveat aside, my understanding of Reviewer 1's comment is that the Reviewer ran a set of simulations which were essentially the same as in the manuscript, except that the Reviewer explicitly excluded the possibility of 90 percent harvest of the stock population.  In the reviewer's simulations, the optimal escapement level is lower in the presence of measurement error, whereas in fig. 5 of the manuscript,  the optimal escapement level is instead higher in the presence of measurement error.

We are afraid that Reviewer 1's concern is mistaken

> Reviewer 1's description of their simulations is brief enough that I would not expect the authors to build their entire revision around an effort to respond to this comment.   If the authors have indeed allowed for 90 percent harvest, however, it is important that they test whether their results are different if they place an upper limit on the fraction of the stock that they allow to be harvested.

When measurements are *uncertain*, it is impossible *not* to allow for a 90% harvest of the stock.  Indeed, this cuts to  the very heart of why measurement uncertainty in *decisions* is so important!  If X is a random number from a log-normal distribution of parameters mu and sigma, how can one choose a number Y to be certain that Y < 90% X?   Y = 0 is the only (non-negative) solution; you cannot be sure that *any* harvest level is guatenteed to be safe.

> I also agree with Reviewer 1's comment that the methods are not that easy to follow.  That said, to my understanding, the escapements in fig. 2 are not constant, and so Reviewer 1's comment that a constant escapement cannot lead to the collapse of a stock seems to me to be irrelevant, but again I may be wrong, and my uncertainty emphasizes the lack of clarity in the writing, especially with respect to the Methods.

The editor is absolutely correct, POMDP solutions do not result in a constant escapement, as we stated.  Unfortunately, Reviewer 1 does not provide any description of the methods they used.

We have revised the

> Page 8-12 in particular are written in a very formal style that I associate with journals that focus on mathematics rather than on biology, with the difficulty problem that much of this text has small grammatical errors that exacerbate the clarity problem.  Meanwhile, both Reviewers argued that the authors have not accurately characterized the literature in this area, and that problem would also have to be solved in a revised version.

We have revised this section significantly to address both grammar and presentation.

The formal definitions and methodology of POMDPs has already been well-documented in the appopriate literature, and there is no reason for us to rehash the all the precise technical details here, which only serves to obscure the most salient concept.  The point here is that there is an immensely well-developed area of theory and tools to address this problem of decision-making under imperfect information.  Though much of that research has come from other fields and is not currently familiar to the readers, we believe we now present enough context to understand the problem.

> In spite of these negative comments, it seems like the overall approach has a great deal of potential, as Reviewer 2 argues, and if Reviewer 1's concerns prove to be minor, then I believe that a revised version would be a substantial contribution.  That said, the complexity of the technical issues is quite high, and so it may be simpler to submit the manuscript to a journal that specializes in mathematical biology or in fisheries management.

Greg Dwyer

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

Reviewer #1:

For complete review, see the attached PDF, which begins as follows:

> I attempted to evaluate the papers overall conclusions by a very simple simulation and these results
seem to challenge the papers conclusions.
I simulated a population with the model the authors describe with both observation and process error
lognormally distributed and using the authors parameters. For a range of escapement goals I calculated
the net present value of using that escapement goal, and compared the consequences when there was
no observation error to the case where the lognormal cv was 0.5 – that is very high observation error.
The one difference I made was I did not let the harvest exceed 90% of the total population size ... this
seems reasonable as I know of few fisheries where 90% of the population could be harvested in a single
year.

As the editor observes, this short discription does not permit us to properly understand what the reviewer has done to obtain
different results, but as the editor already noted, the comments make it clear that it is not analgous to our approach,
since our approach does not assume nor does it find a constant escapement policy to be optimal.  
Moreover, the remark about restricting harvest to 90% of population size is very instructive, as it cuts to the heart of the problem of measurement uncertainty. 
Clearly such a restriction is only possible if the manager has access to perfect observations of the stock size,
an thus any simulation that enforces this restriction is implicitly ignoring the reality of measurement error.

There are only two ways a manager can be certain that harvest does not exceed 90% of stock: (1) don't harvest, or (2) measure the stock size exactly, (or at least, exactly measure some absolute lower bound of possible stock sizes).

As the literature has observed consistently for some time, (e.g. see an eloquent discussion in Clark & Kirkwood 1987) 
if we only consider policies in which accidental population extinction by overharvesting is impossible, 
(e.g. because we assume a priori harvest cannot exceed 90% of the stock), 
there is every reason to suspect this will result in artifically less cautious policies.  If we begin with 
the assumption that there's just no possible way overharvesting could destroy a stock, then the risks of
overharvesting are drastically reduced: net present value need only reflect the loss of revenue while a stock
rebuilds, and not the much greater loss that would result from extinction.  Because even the stochastic theory
such as Reed 1979 made effectively the same assumption (e.g. assumes the stock is measured perfectly before
harvest, so that it would never be rational to attempt a harvest)

> I can’t really understand some of the results in the paper ... for instance I don’t see how a fixed
escapement policy could lead to the collapses shown in Figure 2 – using fixed escapement policies you
can’t have you population continue to decline.

As we discuss above, this is once again fundamental feature of measurement uncertainty -- and again one
that was observed long before our paper and is well established by the literature we discuss at some length.
If a fixed escapement policy is applied to a measurement that is uncertain, the true stock will be somewhat
larger or somewhat smaller than the observed value, and thus the true escapement can be somewhat smaller
or somewhat larger than the 'target' escapement.  Without perfect measurements, it is clearly impossible
to achieve the target escapement exactly.  Overestimates of the stock size can lead to stock collapse

Without perfect measurements, the true escapement will vary under a

The novelty of our result is how to deal with this issue.

> Finally it is certainly not true that fixed escapement policies are commonly used. To my knowledge
almost all marine fish populations are managed with policies more resembling fixed harvest rates with
caveats that target harvest rate declines when stocks drop below a target stock size. Certainly some
salmon stocks are managed with an escapement goal, but they are not in any sense common or the
default in marine fisheries management.

The reviewer is entirely correct about this -- while a rigorous constant escapement policy shuts down
all harvest below B_MSY (the constant escapement level), most fisheries actually continue to harvest
even when the stock drops below the target stock size, though potentially at some decreasing rate,
and do not shut down harvest until the population is significantly below this level, commonly B_MSY / 2, if at all.

As mentioned above, we have now revised all of our analyses to consider these realistic harvest policies,
and have added a substantial discussion addressing the differences between such constant harvest (or constant effort /
fishing mortality) policies and the constant-escapement policies we have hitherto focused upon.  
Because these policies permit harvesting below B_MSY, it is not surprising that we find them even more
suseptible

Moreover, many fisheries do not

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

Reviewer #2:

> In this clearly written paper the authors show that fish population dynamic models that ignore measurement error can result in significant overharvesting, decreased economic returns, and the potential for complete collapse of the stock. The key advance is the use of an algorithm for partially observed Markov decision processes (POMDP) to provide a complete solution rather than relying upon inferences from a deterministic model or a model of optimal control that does not update belief states during each time step.

> The treatment of this issue is remarkably comprehensive with respect to structural uncertainty of the model formulation. In fact, the authors in some way bury what I see as the lead in the Appendix. They include analysis of a population subject to an Allee effect, and produce the same general conclusion as they do via analysis of models with more standard dynamics: the POMDP approach is less likely to cause overharvest and population collapse than alternative approaches, even if the location of the unstable steady state population size (tipping point) is unknown. This finding is very important given the increasingly voluminous literature on thresholds and tipping points, suggesting that proper accounting for measurement uncertainty may be a practical alternative to investing more in reducing such uncertainty in the budget-constrained world of natural resource management.

> As is clear from my summary, this paper is framed for a narrower audience focused on fisheries and stock assessment. To draw in a broader readership, I suggest that the authors motivate the problem they address and interpret their results in the context of the more general ecological, wildlife, and forestry literatures. Specific examples of how measurement uncertainty has been found problematic in fisheries and wildlife assessments would be helpful.

We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here (managing with imperfect information) as well as some other themes (tipping points, management with constrained resources) are far more general. We have now tried to state these implications more clearly.

> The authors mischaracterize the extent to which measurement uncertainty is left untreated in stock assessments. Dichmont et al. 2016 (Fisheries Research) provide a nice review of how some of these measurement uncertainty issues are treated in practice in US stock assessments, and an entry point to the literature. If the authors intend to focus on theoretical models and general solutions for population dynamics, their assertion about failure to incorporate measurement uncertainty completely may be valid but this should be clarified.

We had no intention of suggesting that measurement uncertainty was under-represented in stock assessments; indeed it is precisely the *prevalance* with which stock assessment models account for uncertainty that motivates

> Unlike the classic pretty good yield paper mentioned in the paper (line 440), the authors do not adequately address how conservative harvest is likely to be met with opposition by harvesters and how this issue can be addressed directly by their analysis.

This is an excellent point which we now discuss properly. Our approach highlights that strategies such as Pretty Good Yield, taking 80% of the maximum sustainable yield, are not sufficient solutions to the challenge of measurement uncertainty, since they are still based on fundamentally deterministic premises.

As our analysis shows, unless it is possible to be certain beyond any shadow of a doubt that one has

How do we determine the maximum sustainable yield in the first place?  This number comes from a deterministic stock-recruitment model.  In stochastic models, and even in many deterministic models with age and stage structure, there is no such thing as a maximum sustainable yield, since the optimal yield must fluctuate with the population dynamics.  But at least in those cases one can reasonably assert an average

Stock assessments give us

Measurement errors accumulate.

Pretty Good Yield, taking 80% of the maximum sustainable yield, reflects the same

is defined as a percentage of maximum sustainable yield, which is an essentially deterministic concept.  It is impossible to be sure that

> Similarly, rather than dismiss management strategy evaluation as a less preferable alternative to the optimal control approach taken here, it would be nice to see the authors include a formal contrast. Might the offline POMDP approach be framed this way and results inferred without additional analyses? This would be nice because in the real world, optimal control with annual updating of harvest regulations (the POMDP approach) may not practical. The offline POMDP seems more realistic, perhaps with an online phase implemented such that belief states could be updated in 5-10 year intervals.

 As the reviewer suggests here and as was recognized by Fischer et al (2009) among others, one advantage of the optimal control approach is that it is generally more transparent and unbiased algorithmic way to reach a decision, which to some extent sidesteps the inherently more subjective existing methods, where this issue is acutely felt.  For instance, our value function is a cold-blooded profit equation which places no value in leaving fish in the sea beyond their future return as harvest.  We also ignore more realistic supply-and-demand effects as a conservative assumption, since the decreasing price of higher supply would dampen the incentive for larger harvests (see comparisons in appendix).

> Minor comments:
The analysis related to Fig 5 made me wonder how POMDP compares to alternative frameworks when a stock is increasing or decreasing in abundance. The authors mention this in their interpretation, but I think it could be highlighted more as it seems that the POMDP approach (at least for plaice) allows a more careful tracking of catch to stock size during collapse and rebuilding, a result that would be viewed favorably by fishermen and conservationists alike.

Excellent observation! We have added the following line to emphasize this more clearly:

> Some reorganization may be warranted. Several portions of the Results section describe Methods, which would be nice to signal earlier in the MS (e.g., lines 266-281, lines 298-300 and lines 321-324 could be included in Methods to clarify that the POMDP overestimates measurement uncertainty, etclines 348-351, etc).

> In several places the authors use loose language by referring to marine ecosystem dynamics when in fact they are treating population dynamics.

Thank you, fixed. (TBD)

> The authors discuss volatility of different solutions and would be served well to quantify it. E.g., lines 365-369

Thank you, fixed (TBD)