### 2017-10-07 appendix/appendix.Rmd

---
title: "POMDP comparisons across sigma_m, sigma_g"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: 
  github_document: default
  pdf_document: default
---

```{r aethetics, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r helper_routines, include = FALSE}
det_simulations <- function(x0 = 1/6, Tmax = 20, 
                              B_MSY = K / 2, 
                              F_MSY = r / 2 , 
                              F_PGY = 0.8 * F_MSY){

  msy_policy <- function(x) F_MSY * x
  pgy_policy <- function(x) F_PGY * x
  escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
    if(f(x,0) > B_MSY)
      f(x,0) - B_MSY
    else
      0
  }

  # a simulation function
  det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0
    for(t in 1:(Tmax-1)){
      action[t] <- policy(state[t])
      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
      state[t+1] <- f(state[t], action[t]) 
    }
    data.frame(time = 1:Tmax, state, action, obs)
  }

  # simulate each policy, returning all results in a single unified data frame 
  sims <- 
    list(msy = msy_policy, 
         det = escapement_policy, 
         pgy = pgy_policy) %>% 
    map_df(det_sim, 
           f, x0, Tmax, 
           .id = "method")

  sims
}
```

```{r message=FALSE}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(sarsop)       # the main POMDP package
library(tidyverse)    # for munging and plotting
library(parallel)
```

## Basic deterministic model

```{r}
r <- 0.75
K <- 1

## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
f <- function(x, h){ 
    x + x * r * (1 - x / K) - pmin(x,h)
}
```

Utility (reward) function:

```{r}
reward_fn <- function(x,h) pmin(x,h)
discount <- 1
```

## Calculating MSY

This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.

```{r}
## A generic routine to find stock size (x) which maximizes 
## growth rate (f(x,0) - x, where x_t+1 = f(x_t))
S_star <- optimize(function(x) -f(x,0) + x / discount, 
                   c(0, 2*K))$minimum
```

```{r}
B_MSY <- S_star
MSY <- f(B_MSY,0) - B_MSY

F_MSY <- MSY / B_MSY  
F_PGY = 0.8 * F_MSY
```

### Optional checks:

```{r}
## Make sure these match the known analytic solutions for Graham-Schaefer model:
testthat::expect_equal(B_MSY, K / 2)
testthat::expect_equal(MSY, r * K / 4)
testthat::expect_equal(F_MSY, r / 2)
```

As a basic reference point, simulate these three policies in a purely deterministic world.  Unlike later simulations, here we consider all states an actions exactly (that is, within floating point precision).  Later, states and actions are limited to a discrete set, so solutions can depend on resolution and extent of that discretization.

```{r}
det_simulations(x0 = 1/6, Tmax = 20, B_MSY, F_MSY, F_PGY) %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1) 
```

-----------

##  Introduce a discrete grid

```{r}
## Discretize space
states <- seq(0,2, length=100)
actions <- states
observations <- states
```

We compute the above policies on this grid for later comparison.

```{r}
policies <- list(
  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
```

# POMDP Model

We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values:

```{r}
meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
        mutate(scenario  = as.character(1:length(sigma_m)))
```

```{r}
models <- 
  parallel::mclapply(1:dim(meta)[1], 
           function(i){
  fisheries_matrices(
  states = states,
  actions = actions,
  observed_states = observations,
  reward_fn = reward_fn,
  f = f,
  sigma_g = meta[i,"sigma_g"][[1]],
  sigma_m = meta[i,"sigma_m"][[1]],
  noise = "normal")
})

```

## POMDP solution

The POMDP solution is represented by a collection of alpha-vectors and values, returned in a `*.policyx` file.
Each scenario (parameter combination of `sigma_g`, `sigma_m`, and so forth) results in a separate solution file.

Because this solution is computationally somewhat intensive, we provide

```{r eval=TRUE}
log_dir <- "appendix_alphas" # Store the computed solution files here
dir.create(log_dir)

## POMDP solution (slow, >10,000 seconds per scenario, & memory intensive)
system.time(
  alphas <- 
    parallel::mclapply(1:length(models), 
    function(i){
      log_data <- data.frame(model = "gs", 
                             r = r, 
                             K = K, 
                             sigma_g = meta[i,"sigma_g"][[1]], 
                             sigma_m = meta[i,"sigma_m"][[1]], 
                             noise = "normal")

      sarsop(models[[i]]$transition,
             models[[i]]$observation,
             models[[i]]$reward,
             discount = discount,
             precision = 0.00000002,
             timeout = 10000,
             log_dir = log_dir,
             log_data = log_data)
    })
)

```

We can read the stored solution from the log:

```{r}
meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% mutate(scenario=as.character(1:length(models)))
alphas <- alphas_from_log(meta, log_dir)
```

## Simulating the static policies under uncertainty

```{r static_sims}
Tmax <- 100
x0 <- which.min(abs(K/6 - states))
reps <- 100

static_sims <- 
 map_dfr(models, function(m){
            do_sim <- function(policy) sim_pomdp(
                        m$transition, m$observation, m$reward, discount, 
                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
            map_dfr(policies, do_sim, .id = "method")
          }, .id = "scenario")

```

## Simulating the POMDP policies under uncertiainty

```{r pomdp_sims}
# slower!
unif_prior <- rep(1, length(states)) / length(states)
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
                       reps=reps)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario") %>% left_join(meta)
```

Combine the resulting data frames

```{r}
sims <- bind_rows(static_sims, pomdp_sims) %>%
    left_join(meta) %>%    ## combine output with scenarios
    mutate(state = states[state], action = actions[action]) # index -> value

write_csv(sims, "sims.csv")
```

##  Figure S1

We the results varying over different noise intensities, sigma_g, and sigma_m.  Figure 1 of the main text considers the case of sigma_g = 0.05, sigma_m = 0.1

```{r}
pomdp_sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

## Policy plots

## Historical data estimation

## Historical data POMDP & other solutions

## Historical combined plot