### 2017-12-13 manuscripts/scratch-early-revisions.Rmd
separate out old / scratch drafts from current working draft

Yet fisheries are not managed at their maximum ecological or economic potential.  
Taken as a whole, most ocean fisheries yield a net negative economic return
[@FAO2009], while some 68% of global fisheries are not in good biological condition [@Costello2016].
Nevertheless, fisheries management in both theory and practice are

In a deterministic world at equilibrium (without economic discounting), 
the policies of constant escapement and maximum
sustainable yield look identical: the stock is maintained at a fixed stock
size or biomass, called $B_{MSY}$ in the MSY/fisheries jargon, which is 
equivalent to the escapement $S^*$ in the optimal control jargon and equal
to half the carrying capacity, or $K/2$, in the familiar Graham-Schaefer 
(logistic growth followed by harvest) model. Likewise, the harvest is identical,
occuring after recruitment and bringing the population back to this optimal
level; an annual harvest of $H_{MSY} = rK/4$ in Graham-Schafer model for either policy.
This similarity disguises significant differences in how the policies behave
when the biomass is either above or below $B_{MSY}$.

The economically optimal amount to harvest when the population is below 
$B_{MSY}$ is zero -- no harvest.  This permits the population to return
as rapidly as possible to the stock size at which it generates the maximum
yield, and thus the maximum economic return.  This result holds even when
recruitment is stochastic and thus future stock sizes are uncertain, 
as rigorously proven by @Reed1979.  Perhaps surprisingly, this is not what
fisheries management thinks of as a maximum sustainable yield (MSY) policy,
and does not reflect how fisheries are managed today. Rather, an MSY policy
is intepreted as fishing at the *effort*, $F_{MSY} = H_{MSY} / B_{MSY}$ that would generate the maximum
sustainable yield if the population was at the optimal biomass, $B_{MSY}$,
*regardless* of the current stock size. It is important to realize this policy
is framed in terms of *effort* $F$: for example, the length of the year open to fishing,
or equivalently, the per-capita mortality rate of fish, and not in terms of harvest, $H$.
A population below $B_{MSY}$ that remains harvested at a constant $H_{MSY}$ will eventually be
driven to extinction, but harvested at a constant $H = F_{MSY} \cdot B$ for $B < B_{MSY}$,
the stock will eventually recover to $B_{MSY}$.  Consequently, this simple policy should
converge to the optimal policy in the long run, assuming there is no uncertainty in
recruitment, measurement, or implementation.

The economically optimal policy of constant-escapement also differs from MSY
when stocks are above $B_{MSY}$.  As before, the MSY policy continues to assume
a fixed mortality rate, $F_{MSY}$, which should eventually reduce the stock to
the level of $B_{MSY}$.  Also as before, the optimal policy is not so patient --
to derive maximum yield it seeks to return the population to the $B_{MSY}$ not
*eventually*, but as quickly as possible, and thus will recommend a harvest larger
than $H_{MSY}$ to accomplish this. These differences become more salient in the case
of a population under stochastic growth, since the resulting variation leads
to a population that frequently above or below $B_{MSY}$.  The greater responsiveness
of the constant-escapement policy to the current stock size thus necessarily 
results in both the faster recovery of overexploited stocks and higher economic
returns.

The variation in the harvest required to achieve this constant escapement policy
can seem impractical from a top-down management perspective.  However, it is 
important to realize that the application and implications of this policy are
not purely academic. Economists have for some time observed these inherent 
inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
Approaches such as individual transferable quotas (ITQs) would incentivise
fisheries behavior to converge towards the more dynamic harvests of a constant
escapement policy, as doing so would increase the value of an individual's share.
It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.