### 2017-12-22 appendix/appendix_harvestfirst.Rmd
Appendix example based on f(x - h); recruit before harvest.

---
title: "POMDP comparisons across sigma_m, sigma_g"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: 
  github_document: default
  pdf_document: default
---

```{r aethetics, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r message=FALSE}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(sarsop)       # the main POMDP package
library(tidyverse)    # for munging and plotting
library(parallel)
```

```{r}
options(mc.cores=parallel::detectCores())
```

## Basic deterministic model

```{r}
r <- 0.75
K <- 1

## Unlike classic Graham-Schaefer, this assumes harvest occurs right after assessment, before recruitment
f <- function(x, h){ 
    s <- pmax(x-h, 0) 
    s + s * r * (1 - s / K) 
}
```

Utility (reward) function:

```{r}
reward_fn <- function(x,h) pmin(x,h)
discount <- 0.99
```

## Calculating MSY

This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.

```{r}
## A generic routine to find stock size (x) which maximizes 
## growth rate (f(x,0) - x, where x_t+1 = f(x_t))
S_star <- optimize(function(x) -f(x,0) + x / discount, 
                   c(0, 2*K))$minimum
```

Since we `observe` -> `harvest` -> `recruit`, we would observe the stock at its pre-harvest size, $X_t \sim B_{MSY} + H_{MSY}$.

```{r}
#B_MSY <- S_star     # recruit first, as in classic Graham-Schaefer
B_MSY <- f(S_star,0) # harvest first

#MSY <- f(B_MSY,0) - B_MSY  # recruit first
MSY <- B_MSY - S_star  # harvest first

F_MSY <- MSY / B_MSY  
F_PGY = 0.8 * F_MSY
```

As a basic reference point, simulate these three policies in a purely deterministic world.  Unlike later simulations, here we consider all states an actions exactly (that is, within floating point precision).  Later, states and actions are limited to a discrete set, so solutions can depend on resolution and extent of that discretization.

```{r}

msy_policy <- function(x) F_MSY * x
pgy_policy <- function(x) F_PGY * x

## ASSUMES harvest takes place *before* recruitment, f(x-h), not after.
escapement_policy <- function(x) pmax(x - S_star,0)  
x0 <- K/6
Tmax = 20
det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0

    for(t in 1:(Tmax-1)){
      action[t] <- policy(state[t])
      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
      state[t+1] <- f(state[t], action[t]) 
    }
    data.frame(time = 1:Tmax, state, action, obs)
  }

sims <- 
  list(msy = msy_policy, 
       pgy = pgy_policy,
       det = escapement_policy) %>% 
  map_df(det_sim, 
         f, x0, Tmax, 
         .id = "method") 
sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1)

```

-----------

##  Introduce a discrete grid

```{r}
## Discretize space
states <- seq(0,2, length=100)
actions <- states
observations <- states
```

We compute the above policies on this grid for later comparison.

```{r}

policies <- data.frame(
  det = index(pmax(f(states,0) - S_star,0), actions), # assumes obs->recruit->harvest, f(x_t) - h_t
 # det = index(pmax(states - S_star,0),     actions), # assumes obs->harvest->recruit, f(x_t - h_t)
  msy = index(states * F_MSY,               actions),
  pgy = index(states * F_PGY,               actions))
```

# POMDP Model

We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values:

```{r}
meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
        mutate(scenario  = as.character(1:length(sigma_m)))
log_dir <- "appendix_alphas" # Store the computed solution files here
meta
```

```{r}
models <- 
  parallel::mclapply(1:dim(meta)[1], 
           function(i){
  fisheries_matrices(
  states = states,
  actions = actions,
  observed_states = observations,
  reward_fn = reward_fn,
  f = f,
  sigma_g = meta[i,"sigma_g"][[1]],
  sigma_m = meta[i,"sigma_m"][[1]],
  noise = "normal")
})

```

## POMDP solution

The POMDP solution is represented by a collection of alpha-vectors and values, returned in a `*.policyx` file.
Each scenario (parameter combination of `sigma_g`, `sigma_m`, and so forth) results in a separate solution file.

Because this solution is computationally somewhat intensive, be sure to have ~ 4 GB RAM per core if running the 9 models in parallel.  Alternately, readers can skip the evaluation of this code chunk and read the cached solution from the `policyx` file using the `*_from_log` functions that follow:

```{r eval=FALSE}
dir.create(log_dir)

## POMDP solution (slow, >10,000 seconds per scenario, & memory intensive)
system.time(
  alphas <- 
    parallel::mclapply(1:length(models), 
    function(i){
      log_data <- data.frame(model = "gs", 
                             r = r, 
                             K = K, 
                             sigma_g = meta[i,"sigma_g"][[1]], 
                             sigma_m = meta[i,"sigma_m"][[1]], 
                             noise = "normal",
                             scenario = meta[i, "scenario"][[1]])

      sarsop(models[[i]]$transition,
             models[[i]]$observation,
             models[[i]]$reward,
             discount = discount,
             precision = 0.00000002,
             timeout = 10000,
             log_dir = log_dir,
             log_data = log_data)
    })
)

```

We can read the stored solution from the log:

```{r}
meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% left_join(meta) %>% arrange(scenario)
alphas <- alphas_from_log(meta, log_dir)
```

## Simulating the static policies under uncertainty

```{r static_sims}
Tmax <- 100
x0 <- which.min(abs(K/6 - states))
reps <- 100

static_sims <- 
 map_dfr(models, function(m){
            do_sim <- function(policy) sim_pomdp(
                        m$transition, m$observation, m$reward, discount, 
                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
            map_dfr(policies, do_sim, .id = "method")
          }, .id = "scenario")

```

## Simulating the POMDP policies under uncertiainty

```{r pomdp_sims}
# slower!
unif_prior <- rep(1, length(states)) / length(states)
pomdp_sims <- 
  map2_dfr(models, alphas, function(.x, .y){
             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
                       reps = reps)$df %>% 
              mutate(method = "pomdp") # include a column labeling method
           },
           .id = "scenario")
```

Combine the resulting data frames

```{r}
sims <- bind_rows(static_sims, pomdp_sims) %>%
    left_join(meta) %>%    ## combine output with scenarios
    mutate(state = states[state], action = actions[action]) # index -> value

write_csv(sims, "appendix_files/sims.csv")
```

##  Figure S1

We the results varying over different noise intensities, sigma_g, and sigma_m.  Figure 1 of the main text considers the case of sigma_g = 0.05, sigma_m = 0.1

```{r}
sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

## Policy plots

Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (deterministic) recruitment would land us right at it $B_{MSY}$.  This corresponds to constant escapement.  The discrete grid makes these appear slightly stepped.

Note that the deterministic solution crosses the MSY solution at an observed value of $B_{MSY}$ (i.e. $K/2 = 0.5$).  PGY harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, PGY and MSY solutions never go to zero.

```{r}
policy_table <- tibble(state = 1:length(states)) %>%
  bind_cols(policies) %>%
  gather(policy, harvest, -state) %>%
  mutate(harvest = actions[harvest], state = states[state])

write_csv(policy_table, "appendix_files/policy_table.csv")
```

```{r}
policy_table %>% 
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.

```{r}
policy_table  %>%
  mutate(escapement = f(state,harvest))  %>% 
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line() + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))

```

### Adding the POMDP policy solution