### 2018-07-30 reviews/README.Rmd
reviews

---
title: "README.Rmd"
author: "Carl Boettiger"
date: "12/13/2017"
output: github_document
---

# README

Reviews are copyright of the reviewers and cannot be shared without permission.  To keep everything in one place, review material is stored here in encrypted files.

```{r}
library(gpg)
keyring <- gpg_list_keys(secret = TRUE)
cboettig <- keyring[1, "id"]
```

Encrypt `.gitignore`'d plain text, allowing only specified users to decrypt:

```{r}
## letter 1
letter1 <- "review-reply.Rmd"
gpg_letter1 <- "review-reply.Rmd.gpg"

msg <- gpg_encrypt(letter1, receiver = cboettig, signer = cboettig)
writeLines(msg, gpg_letter1)
```

Unencrypt committed gpg versions:

```{r}
writeLines(gpg_decrypt(gpg_letter1), letter1)
```

Encrypt letter 2

```{r}
letter2 <- "review-reply-2.Rmd"
gpg_letter2 <- "review-reply-2.Rmd.gpg"

msg <- gpg_encrypt(letter2, receiver = cboettig, signer = cboettig)
writeLines(msg, gpg_letter2)
```

Unencrypt committed gpg versions:

```{r}
writeLines(gpg_decrypt(gpg_letter2), letter2)
```